GVN-Hoist: Hoisting Computations from Branches

### 2016-08-09 hoist.tex
policies and partition algo

\section{Code hoisting policies}
The amount of hoisting depends on the whether we collect information about all
the instructions before finding candidates (optimistic).

In the pessimistic approach, the basic blocks are traversed in the inverse
depth-first order, computing the GVN of instructions as they come by. The GVN of
sibling branches are compared for equality. Once such a candidate is found, it
is hoisted in the common dominator. All the leaf nodes are visited before the
non-leaf nodes (bottom-up) because instructions are hoisted upwards. The
pessimistic algorithm is fast, results in fewer spills but hoists very less
instructions. We have implemented optimistic approach because it is more general
and can be tuned down to closely mimic pessimistic approach by changing a flag.

It also depends on the generality of the GVN algorithm, however, that analysis
is beyond the scope of this paper.

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We collect the GVN of all the instructions in the function and iterate on the
list of instructions having identical GVNs. After that we find the common
dominator dominating all such identical computations and perform legality checks
(Section-\ref{subsec:legality}). Many times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints e.g.,
intersecting side-effects or, profitability constraints e.g., hoisting point too
far. In those cases, this algorithmn would partition the list of identical
computations and partially hoist a set of computations to their common
dominator. The partition algorithm is described as follows:

\subsubsection{Partition the list of hoisting candidates to maximize hoisting}
\label{subsec:partition}
The set of candidates are sorted in the order of their DFS-in numbers. So that
candidates close to each other in the list are also closer (??) in the CFG.

In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximise the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their DFS-in numbers,
we make sure that candidates closer in the list have their common dominator
nearby in cases when there are no fully redundant instructions.

@@

>>>>

\section{Code hoisting policies}
The amount of hoisting depends on whether we collect GVN of instructions
on-demand (pessimistic) or, before finding candidates (optimistic).It also
depends on the generality of the GVN algorithm, however, that analysis is beyond
the scope of this paper.

\subsection{Hoisting bottom up with pessimistic approach}
\label{subsec:pessimistic}
In the pessimistic approach, the basic blocks are traversed in the inverse
depth-first order, computing the GVN of instructions as they come by. The GVN of
sibling branches are compared for equality. Once such a candidate is found, it
is hoisted in the common dominator. All the leaf nodes are visited before the
non-leaf nodes (bottom-up) because instructions are hoisted upwards.

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We collect the GVN of all the instructions in the function and iterate on the
list of instructions having identical GVNs. After that we find the common
dominator dominating all such identical computations and perform legality checks
(Section-\ref{subsec:legality}). Many times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints e.g.,
intersecting side-effects or, profitability constraints e.g., hoisting point too
far. In those cases, this algorithmn would partition the list of identical
instructions into subsets which can be partially hoisted to their respective
common dominators. The partition algorithm is described as follows:

\subsubsection{Partition the list of hoisting candidates to maximize hoisting}
\label{subsec:partition}
In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximise the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their DFS-in numbers,
we make sure that candidates closer in the list have their common dominator
nearby in cases when there are no fully redundant instructions.

@@

The pessimistic algorithm is fast, results in fewer spills but hoists very less
instructions. We have implemented optimistic approach because it is more general
and can be tuned down to closely mimic pessimistic approach by changing a flag.
